---
title: "Week 13 Main Assignment"
author: "Avi Kodali"
output: html_notebook
---

```{r setup, include = FALSE}
library(tidyverse)
library(scales)
library(tabulizer)
```

## Problem 1
* Problem 1 is the bad plot assignment, to be submitted separately.

## Problem 2
* The chunk below will download a pdf file, from the NJ courts system, that gives the number of non-English language speakers in NJ. (The data actually comes from the ACS, and is based on a pair of questions, the first asking if the person speaks a language other than English at home and if so, what that language is.) 
* Although the data is available more easily elsewhere, this file gives some nice practice in getting data from a pdf file.



```{r file download}
if (!file.exists("NJ-languages.pdf")){
download.file("https://www.njcourts.gov/public/assets/langSrvcs/ranklang2016-2019.pdf?c=C3N", "NJ-languages.pdf")
}
```
* In the code chunk below, use `extract_tables()` with the argument `output = "data.frame"` to extract the data from the pdf file. Name the resulting object `prob2_lst`.
* Don't worry about warnings about illegal operations.

```{r prob 2a}
prob2_1st <- extract_tables("NJ-languages.pdf", output = "data.frame")
```

* You can see from the code chunk below that the columns in the first and second pages have been interpreted somewhat differently.

```{r}
glimpse(prob2_1st[[1]])
glimpse(prob2_1st[[2]])
```

* The code below shows that there is another problem with reading the second page:

```{r}
prob2_1st[[2]] %>% 
  slice(7:8)
```

* For this assignment, however, we will work only with the data frame that is the first element of `prob2_lst`. 
* Assign that first element to the object `prob2_tbl`, remove the column `Language`, and rename the column `X` to `Language`.

```{r prob 2b}
prob2_tbl <- prob2_1st %>% 
  .[[1]] %>% 
  select(-Language) %>% 
  rename(Language = X)
```

* You will notice that the `Number.of.Speakers` variable is of type character, not number---that happened because of the commas. Please take care of that issue in the code chunk below.

```{r}
prob2_tbl <- prob2_tbl %>% 
  mutate(Number.of.Speakers = parse_number(Number.of.Speakers))
```

* Finally, please make a bar plot of the top 10 languages spoken at home in NJ, after English, with the bars in decreasing order.
  - Use "Number of Speakers" as the label for one axis, and omit the label for the other.
  - Make sure the tick labels for Number of Speakers look nice by making sure they have commas in them.
  - The language names are rather long, so place them at a 45 degree angle.

```{r}
prob2_tbl %>% 
  arrange(desc(Number.of.Speakers)) %>% 
  head(10) %>% 
  ggplot(aes(fct_rev(fct_reorder(Language, Number.of.Speakers)), y = Number.of.Speakers)) +
  geom_col() +
  theme_bw() +
  scale_y_continuous(name="Number of Speakers", labels = comma) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x.bottom = element_text(angle = 45, vjust = 0.5),
    plot.title = element_text(hjust = 0.5)) + 
  ggtitle("Number of Speakers for the Top 10 Languages in NJ")
```


## Problem 3

* The code chunk below establishes a connect to a database that contains tables of gdp_per_cap, health_spending, infant_mortality, population corresponding exactly to the Problem 2 of the Week 12 assignment.

```{r}
con <- DBI::dbConnect(RSQLite::SQLite(), dbname = "w13-assignment.sqlite")
```

* Using SQL code, please create a table by starting with the `population` table and restricting to countries that had a population of at least 1,000,000 in the year 2000.

```{sql problem 3a, connection = con, output.var = "prob_3a"}
SELECT *
  FROM population
  WHERE "2000.0" >= 1000000;
```

* The code chunk below should show that the table resulting from your code had a dimension of 166, 82.
```{r problem 3a check}
dim(prob_3a)
```


* Now create a table that consists of the 5 countries with the largest gdp per capita in 2000 among countries that had a population of at least 1,000,000 in the year 2000. You will want to reuse your code from chunk problem 3a.
* Hint: Recall that SQLite does not have a keyword for semi join, but you can follow the pattern from slides 16 and 17 of Week 12 Class 1.


```{sql problem 3b, connection = con, output.var = "prob_3b"}
SELECT * 
  FROM gdp_per_cap
LEFT JOIN ( SELECT country
              FROM population
              WHERE "2000.0" >= 1000000
          ) AS temp
  ON gdp_per_cap.country = temp.country
  WHERE temp.country IS NOT NULL
ORDER BY "2000" DESC
LIMIT 5;
```

* If you were successful, the code chunk below should show
```
Norway
Japan
Switzerland
United States
United Arab Emirates
```

```{r problem 3b check}
prob_3b[, 1] %>%
  cat(sep = "\n")
```

## Finishing up

* When you are done, be sure that you can knit your file to confirm that your code runs start to finish without errors. Besides modifying the code chunks, make sure you have replaced YOUR CODE HERE, YOUR NAME HERE, and YOUR ANSWER HERE appropriately. Be sure to sign the honor pledge below.

* Finally, be sure that you have saved the final version of your file in RStudio Cloud (under the File menu of RStudio Cloud, not the File menu of your browser). Download the file to your own computer, and from there, upload the file to Canvas. There is no need to rename it---Canvas will add your name to it when I download it.

On my honor, I have neither received nor given any unauthorized assistance on this assignment.
Avi Kodali
