---
title: "Week 1 Main Assignment"
author: "Avi Kodali"
output: html_document
---

## Instructions

*You may use whatever resources you like for the following assignment, including working with others.* The work submitted should be your own, however, and not simply copied.


## Set up
First load some required packages. You can run the entire chunk by clicking on the green triangle below.


```{r setup, include=FALSE}
library(tidyverse)
library(babynames)
```


## Problem 1

The code chunk below is the same as the first chunk in `02-first-example.Rmd`. Please modify it (including the title) to produce a plot that will answer the following question: If you are going to meet a woman named Susan and you know nothing about her, what is your best guess as to when she was born? 


```{r, problem-1}
babynames %>% 
 filter(name == "Susan", sex == "F") %>%
 ggplot(aes(year, n)) +
 geom_line() +
 ggtitle("Popularity of Name Susan") 
```
 
Based on the plot, a woman named Susan was more likely born in 1955 than any other year. (Your answer should be correct within 5 years.)
 
## Problem 2

The code chunk below is the same as the first chunk in `02-first-example.Rmd`. Please modify it (including the title) to produce a plot that will answer the following question: Is a boy born in 2000 more likely to be named Adam or Bryce? 

```{r, problem-2}
babynames %>% 
 filter(name %in% c("Adam","Bryce"), sex == "M") %>%
 ggplot(aes(year, n, color = name)) +
 geom_line() +
 ggtitle("Popularity of Name Adam vs Popularity of Name Bryce") +
 coord_cartesian(xlim = c(1999, 2001))
```

Based on the plot, Adam is the more common of the two names for boys born in 2000.


 
 
## Finishing up

When you are done, be sure to select "Run All" from the "Run" menu (top right of the source code pane) to confirm that your code runs start to finish without errors. Besides modifying the two code chunks, make sure you have replaced YOUR NAME HERE and YOUR ANSWER HERE appropriately. Be sure to sign the honor pledge below.

Finally, be sure that you have saved your file in RStudio Cloud (under the File menu of RStudio Cloud, not the File menu of your browser). Then, in the Files pane, click the checkbox by the file named `03-main-assignment.Rmd` and choose "More" from the pane's menu, followed by "Export", to download the file to your own computer. From there, upload the file to Canvas. There is no need to rename it---Canvas will add your name to it when I download it.

The downloading procedure can be a little confusing the first time---figuring it out is part of the assignment---so don't wait until the last minute.


## Honor pledge

On my honor, I have neither received nor given any unauthorized assistance on this quiz.
Avi Kodali
